working title Compatibility system and stygma size of pollen recipient as main predictors of heterospecific pollen effect

### 2018-12-10 Manuscript_draft/Manuscript_draft.Rmd
Changes for meeting...

Heterospecific pollen reduced seet set signifcatively in xxx from xxx pairwise species combinations. Being greater in xxx and then in xxx. I have tested that with linear mixed effect models (add this to methods). Add boxplots here...

Mantel test gave significant values p<0.05 for both types of evolutive distances considered, with r coefficients of 0.29 and 0.25 for the distances of ITS and RBCL markers respectively. Moreover, Mantel test gave a significant correlation between stigma width and stigma type with xxxx. Trait correlations were also explored with ... and we found that...

>>>>

Heterospecific pollen reduced seet set signifcatively with the 50-50% heterospecific pollen treatments for 65% of the paiwise interactions p<0.05. Across families we found a very similar effect but when species where look at species level they respond differently even within the same family, for instance for two species of the Brassicaceae family Brassica oleracea and Eruca sativa we found very contrasting effects of foreign pollen where for the first one, all donors reduce seedt set significatively and for the second, just two species did out of nine. The treatments of 100% heterospecific pollen ... didn't seed except...

Mantel test indicates that a possible correlation exist between heterospecific pollen effect and the evolutive distances, for  ITS and RBCL markers we had r coefficients of 0.29 and 0.25 respectively p<0.05. Moreover, Mantel test indicates that also a possible correlation between stigma width and stigma type exist. Trait correlations were also explored with ... and we found that...

### 2018-12-10 Manuscript_draft/Manuscript_draft.tex
Changes for meeting...

Interestingly, incompatibility system seems to play an important role in
foreign pollen effect where species that are self incompatible would
have stronger barriers towards heterospecific pollen than self
compatible species Ashman and Arceo-Gómez (2013). The type of
incompatibility, sporophytic or gametophytic is related with the place
of pollen recognition where the former take place at the sitgma level
and the latter occurs within the style, this last late acting pollen
recognition mechanism is associated with greater negative effect Barrett
(1988). Remarkably, there is a great variability in mating systems
across populations Whitehead et al. (2018) and therefore predict an
effect of foreign pollen is a bit obscured by the variability within
species, however species that are strong selfers or strong outcrossers
have less variablity in mating systems and predictions of effect could
be more realistic (see figure 1 from Whitehead et al. (2018)). Moreover,
other traits such as number of pollen grains per flower and number of
ovules have been tradittionally associated with the type of
incompatibility system where species with higher pollen ovule ratios are
predicted to be xenogamous and species with low pollen ovule ratios
autogamous (REF). Selfer species would have a reduction of herkogamy
(REF), less pollen production per ovule (REF) and therefore less likely
to contribute in exporting pollen in the community. Other morphological
traits, like stigma size can be determinant for the total pollen quatity
that a stigma can receive and therefore related to do that pollen size
would also play an important role. Example with pollen here.

@@

The great environmental variability in natural systems and complexity of
floral structures make heterospecific pollination studies a daunting
task. Moreover, variation in sampling effort have been shown to be
determinant to characterize pollen transfer interactions Arceo-Gómez et
al. (2018). Although plant-pollinator network and pollen network studies
can give a first picture of the importance of foreign pollen is
necessary to address how its effect is shaped with both traits and
relatedness of the species. For this reason, in this study we have
created an artificial co-flowering community with 10 species belonging
to three different families where we try to test the following
questions: 1) Does heterospecific pollen reduce seed set, if so, 2) Does
heterospecific pollen effect depend on the relatedness of the species,
3) Does heterospecific pollen effect depend on any floral trait?

@@

Foreign pollen effect was studied through two different treatments, one
with 50\% conspecific pollen and 50\% heterospecific pollen and a second
one with 100\% foreign pollen (N=10). Therefore, 180 different
combinations were perform with N=10. Seed set was the proxy of effect
for all our teratments. Moreover, hand cross pollination, hand self
pollination, apomixis (bagged emasculated flowers) and natural selfing
were tested (N=10). Flowers were emasculated the day prior anthesis and
hand pollinated next day with a toothpick. HaNd-pollination was realized
with 3-4 gentle touches on the stigma surface. The mixes of pollen were
performed on an eppendorf based on the pollen counts maded with Neubaeur
chamber (each anther was counted 4 times for 20 different anthers per
species). In order to confirm that the treatments applied were 50-50\%
pollen we counted for each species the total pollen stigmatic load from
a donor of each family (N=3).

@@

The traits measured for each species were pollen per anther, number of
ovules, stigma width and length and stigmatic area, style width and
length, ovary width and length. Moreover stigma type was tested. Pollen
was counted for 20 anthers of each species with 4 replicates per sample
with an hemocytometer. Previously anthers were squashed on a known
solution with the pippete tip and homogeneize with a vortex for 30
seconds. Ovule number was counted with the help of an stereomicroscope
and a small grid over a petri dish from 15 randomly selected flowers.
The different morphometrical traits were measured with XXXX. Levels of
self incompatibility were estimated by dividing the the fruit set of
hand self pollination by hand cross pollination. Moreover, two types of
evolutive distances were calculated with MEGA7 for two kinds of markers:
1) Internal transcribed spacer (ITS) and 2) ribulose-bisphosphate
carboxylase (RBCL)

@@

We used the statistical language \texttt{R} (R Core Team 2018) for all
our analyses. These were implemented in dynamic rmarkdown documents
using \texttt{knitr} (Xie 2014, 2015, 2018) and \texttt{rmarkdown}
(Allaire et al. 2018) packages. To test the effect of heterospecific
pollen, we substracted the to seed set of hand cross pollination the
seed set of the heterospecific pollen treatments. Therefore, small
values mean low effect and viceversa. To be able to compare among
species seed set was previosly scaled for all species. In order to see
relations between hetereospecific pollen effect and traits we performed
Mantel test between the matrix of effect and the distance matrix of each
trait (euclidean distances). Moreover, Mantel test was also conducted
between heterospecific pollen effect and phylogenetic distance.

@@

Heterospecific pollen produced fitness for xxx cases.

@@

\hypertarget{ref-Allaire_2018}{}
Allaire, J., Y. Xie, J. McPherson, J. Luraschi, K. Ushey, A. Atkins, H.
Wickham, J. Cheng, and W. Chang. 2018. Rmarkdown: Dynamic documents for
r.

@@

@@

\hypertarget{ref-Xie_2014}{}
Xie, Y. 2014. Knitr: A comprehensive tool for reproducible research in
R. \emph{in} V. Stodden, F. Leisch, and R. D. Peng, editors.
Implementing reproducible computational research. Chapman; Hall/CRC.

\hypertarget{ref-Xie_2015}{}
Xie, Y. 2015. Dynamic documents with R and knitr. 2nd editions. Chapman;
Hall/CRC, Boca Raton, Florida.

\hypertarget{ref-Xie_2018}{}
Xie, Y. 2018. Knitr: A general-purpose package for dynamic report
generation in r.

>>>>

Interestingly, incompatibility system seems to play an important role in
foreign pollen effect where species that are self incompatible would
have stronger barriers towards heterospecific pollen than self
compatible species Ashman and Arceo-Gómez (2013). The type of
incompatibility, sporophytic or gametophytic is related with the place
of pollen recognition where the former take place at the sitgma level
and the latter occurs within the style, this last late acting pollen
recognition mechanism is associated with greater negative effect Barrett
(1988). Remarkably, there is a great variability in mating systems
across populations Whitehead et al. (2018) and therefore predict an
effect of foreign pollen is a bit obscured by the variability within
species, however species that are strong selfers or strong outcrossers
have less variablity in mating systems and predictions of effect could
be more realistic (see figure 1 from Whitehead et al. (2018)). Moreover,
other traits such as number of pollen grains per flower and number of
ovules have been tradittionally associated with the type of
incompatibility system where species with higher pollen ovule ratios are
predicted to be xenogamous and species with low pollen ovule ratios
autogamous (REF). Selfer species would have a reduction of herkogamy
(REF) and less pollen production per ovule (REF) which can be
interpretated as a reduction of pollen expoorted into the community.
Other morphological traits, like stigma size can be determinant for the
total pollen quatity that a stigma can receive and therefore related to
do that pollen size would also play an important role. Example with
pollen here.

@@

The great environmental variability in natural systems and complexity of
floral structures make heterospecific pollination studies a daunting
task. Moreover, variation in sampling effort have been shown to be
determinant to characterize pollen transfer interactions Arceo-Gómez et
al. (2018). Although plant-pollinator network and pollen network studies
can give a first picture of the importance of foreign pollen is
necessary to address how its effect is shaped with both traits and
relatedness of the species. For this reason, in this study we have
created an artificial co-flowering community with 10 species belonging
to three different families with different traits where we try to test
the following questions: 1) Does heterospecific pollen reduce seed set,
if so, 2) Does heterospecific pollen effect depend on any floral trait?
3) Does heterospecific pollen effect depend on the relatedness of the
species.

@@

Foreign pollen effect was studied through two different treatments, one
with 50\% conspecific pollen and 50\% heterospecific pollen and a second
one with 100\% foreign pollen (N=10). Therefore, 180 different
combinations were perform with N=10. Seed set was the proxy of effect
for all our treatments. Moreover, hand cross pollination, hand self
pollination, apomixis (bagged emasculated flowers) and natural selfing
were tested for each species (N=10). Flowers were emasculated the day
prior anthesis and hand pollinated next day with a toothpick.
Hand-pollination was conducted with 3-4 gentle touches on the stigma
surface. The mixes of pollen were realized on an eppendorf based on the
pollen counts maded with Neubaeur chamber (each anther was counted 4
times for 20 different anthers per species). In order to confirm that
the treatments applied were 50-50 percent pollen, for each focal species
the total stigmatic load of pollen was counted from one donor of each
family (N=3).

@@

The traits measured for each species were pollen per anther, number of
ovules, stigma width and length and stigmatic area, style width and
length, ovary width and length. Moreover stigma type was tested. Pollen
was counted for 20 anthers of each species with 4 replicates per sample
with an hemocytometer. Previously anthers were squashed on a known
solution with the pippete tip and homogeneize with a vortex for 30
seconds. Ovule number was counted with the help of an stereomicroscope
and a small grid over a petri dish from 15 randomly selected flowers.
The different morphometrical traits were measured with a digital
stereomicrospe. Levels of self incompatibility were estimated by
dividing the the fruit set of hand self pollination by hand cross
pollination Lloyd and Schoen (1992).

@@

We used the statistical language \texttt{R} (R Core Team 2018) for all
our analyses. To test the effect of heterospecific pollen, we
substracted to the seed set of hand cross pollination the seed set of
heterospecific pollen treatments. Therefore, small values mean low
effect and viceversa. To be able to compare among species, seed set was
previosly scaled with mean 0 and standard deviation of 1. In order to
see correlations between hetereospecific pollen effect and traits we
performed Mantel test between the matrix of effect and the distance
matrix of each trait (euclidean distances). Moreover, Mantel test was
also conducted between heterospecific pollen effect and the square root
of the matrix of phylogenetic distance due to improvement in the
statistical power (Letten \& Cornwell 2014). We explored also the the
relations between traits and heterospecific pollen effect through
generalized mixed models where the response variable was heterospecific
pollen effect, the independent variable the different traits and the
random effects the different treatments per species. Moreover, pairwise
evolutive distances were calculated with MEGA7 for two kinds of markers:
1) Internal transcribed spacer (ITS) and 2) ribulose-bisphosphate
carboxylase (RBCL). The sequences of interest were downloaded from NCBI
GenBank and the phylogenetic tree constructed by maximum likelihood with
MEGA7.

Phylogenetic signal of traits?

@@

Heterospecific pollen reduced seet set signifcatively with the 50-50\%
heterospecific pollen treatments for 65\% of the paiwise interactions
p\textless{}0.05. Across families we found a very similar effect but
when species where look at species level they respond differently even
within the same family, for instance for two species of the Brassicaceae
family Brassica oleracea and Eruca sativa we found very contrasting
effects of foreign pollen where for the first one, all donors reduce
seedt set significatively and for the second, just two species did out
of nine. The treatments of 100\% heterospecific pollen \ldots{} didn't
seed except\ldots{}

Mantel test indicates that a possible correlation exist between
heterospecific pollen effect and the evolutive distances, for ITS and
RBCL markers we had r coefficients of 0.29 and 0.25 respectively
p\textless{}0.05. Moreover, Mantel test indicates that also a possible
correlation between stigma width and stigma type exist. Trait
correlations were also explored with \ldots{} and we found that\ldots{}

Fix mantel test selfing rates and change it for compatibility
index\ldots{}

Fix this to GLMM? Yep I have to\ldots{}

@@

@@

\hypertarget{ref-lloyd1992}{}
Lloyd, D. G., and D. J. Schoen. 1992. Self-and cross-fertilization in
plants. i. functional dimensions. International Journal of Plant
Sciences 153:358--369.

@@
